---
title: "Reproduction of Palmer Penguins Dataset Analysis"
subtitle: 'MSc in Statistics 2025/26, Imperial College London'
author: "Kevin O'Loughlin, CID: 06059955"
format:
  html:
    toc: true
    highlight: tango
    self-contained: true
    df-print: paged
  pdf: default
format-links: false
bibliography: "MSc_Statistics_coursework_template_quarto.bib" 
---

<style type="text/css">
h1{
  font-size: 24pt;
}
h2{
  font-size: 18pt;
}
body{
  font-size: 12pt;
}
</style>

```{r setup, include = FALSE, tidy=TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
include_solutions <- TRUE
```
```{r setup2, include=FALSE, tidy=TRUE}
require(rmarkdown)
require(knitr)
require(kableExtra)
# Put any library imports and other preamble here.
```


In this document, we will perform analysis on the _penguins_ dataset accessed from the _palmerpenguins_ package (accessed below).


```{r}
#| warning: false
library(palmerpenguins)

# use :: to access data or functions from within a package 
penguins <- palmerpenguins::penguins
```

```{r, message=FALSE}
#| label: sex-counts
#| tbl-cap: "Number of Penguins by Sex"
sex_counts <- table(penguins$sex)
knitr::kable(sex_counts, col.names  = c("Sex", "Count"))
```

@tbl-sex-counts Shows the number of male and female penguins in the dataset.


```{r, echo=FALSE}
#| layout-ncol: 2
#| layout-nrow: 2
#| label: box-whisker-plots
#| fig-cap: "Distributions of Features by Sex"
#| fig-subcap: 
#|   - "Bill Length Distribution"
#|   - "Bill Depth Distribution"
#|   - "Flipper Length Distribution"
#|   - "Body Mass Distribution"

vars <- c("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g")


for (v in vars) {
  plot(penguins$sex, penguins[[v]],
       xlab = "",
       ylab = gsub("_", " ", v),
       col = c(6, 3))
  abline(h = mean(penguins[[v]], na.rm = TRUE), col = 2, lwd = 2)
}
```

@fig-box-whisker-plots demonstrates the distribution of various features for both male and female penguins.

@penguins proposes a method for identifying the gender Otago blue penguins from bill measurements, suggesting that the variations by gender seen here may reflect real patterns.

```{r}
#| label: bill-length-body-mass-scatter
#| fig-cap: "Penguin Bill Length vs Body Mass"
plot(
 x = penguins$bill_length_mm,
 y = penguins$body_mass_g,
 col = penguins$sex,
 pch = 15 + as.numeric(penguins$sex),
 xlab = "Bill Length (mm)",
 ylab = "Body Mass (g)",
 bty = "n")
 legend(
 "bottomright",
 legend = c("Female", "Male"),
 pch = 16:18,
 col = 1:3,
 bty = "n")
```

@fig-bill-length-body-mass-scatter analyses the relationship between the Bill Length and Body Mass for each sex.



### References

::: {#refs}
:::